Pesticides in small agricultural streams in Germany

### 2016-08-19 article.tex
Revert "restructured (moved latex file to separate folder)"
This reverts commit 3f4daa0bd542b0f2452e3dde6c7592b911bdce4a.

\citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
However, this reflects only a small part of data from national scale monitoring programs.
National monitoring programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
These monitoring programs produce large amounts of data, which possibly can also be used to answer other questions.
In Germany monitoring programs are setup independently by the federal states in compliance with the Water Framework Directive \citep{quevauviller_water_2008} and additional state specific needs.
However, currently there is no curated national-wide compilation of this data available.
\citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
They found that insecticides that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.

Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonisation of polluted streams \citep{liess_analyzing_2005}.
At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
Although small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007} relatively little is know about their chemical and ecological status.

@@

We homogenized and unified these datasets into a common database.
We implemented a robust and transparent data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
An overview of the data cleaning process is provided in the supplemental materials.  
To assess whether samples were taken during potential rainfall events we performed spatio-temporal intersection of sampling events with daily precipitation data from the sampling date and the day before. \citep{rauthe_central_2013}.

@@

\subsection{Pollution of small agricultural streams}
Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
%EQS
Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:fig5}A).
% RAC
Neonicotinoid Insecticides and Chlropyrifos showed highest risk quotients.
For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
In 15\% of all samples (n = 6377) risc quotients higher then 1 were observed.
Highest RQ were observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
% TU
In 33\% of the samples no pesticides were detected. 
The mean $log(TU_{max})$ for detects was -4.6.
2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
We found a high correlation of $log(TU_{max})$ and $TU_{sum}$ \todo{ausrechnen und r angeben}, with a maximum higher value of $TU_{sum}$ xx units.

%Mixtures
In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:fig5}D).

@@

% current problems in monitoring and possible sollutions
A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analysed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.

@@

% RAC
\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances organophosphate insecticides. 
Our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
This can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here \todo{zahlen neonics} and shows that this class may pose a high risk to freshwaters \todo{anderes Wort}.
Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality. 
This is also reflected in the different sample sizes (1,566 vs. 2,918,604 measurements \todo{use number for substances with RAC}) and the high number of non-detects in the monitoring data.

@@

% TU
xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshhold above biological effects have been shown \todo{add citation schaefer}. 
The high correlation of TU and the relatively low exceedance of $TU_sum$ reveals that in compound mixtures occurring in the field there most only one compound dominating the toxicity.

% Mixtures
Nevertheless, most pesticides did not occur individually but in mixtures.

>>>>

\citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
However, this study reflected only a small part of data that is available from national monitoring programs.
These programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
Large amounts of data are generated, which possibly can also be used to answer other questions.
In Germany monitoring programs are setup independently by the federal states in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
However, currently there is no curated national-wide compilation of this data available.

\citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
They found that insecticides that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.
Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonisation of polluted streams \citep{liess_analyzing_2005}.
At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
Although small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, relatively little is know about their chemical and ecological status.

@@

We homogenized and unified these datasets into a common database.
We implemented a robust and reproducible data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
An overview of the data cleaning process is provided in the supplemental materials.  
To assess whether samples were taken during potential rainfall events we performed spatio-temporal intersection of sampling events with daily precipitation data from the sampling date and the day before. \citep{rauthe_central_2013}.

@@

\subsection{Pollution of small agricultural streams}
Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
%EQS
Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:fig5}A).
% RAC
Neonicotinoid Insecticides and Chlropyrifos showed highest risk quotients.
For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
In 15\% of all samples (n = 6377) risc quotients higher then 1 were observed.
Highest RQ were observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
% TU
In 33\% of the samples no pesticides were detected. 
The mean $log(TU_{max})$ for detects was -4.6.
2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
We found a high correlation of $log(TU_{max})$ and $TU_{sum}$ \todo{ausrechnen und r angeben}, with a maximum higher value of $TU_{sum}$ xx units.
%Mixtures
In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:fig5}D).

@@

% current problems in monitoring and possible solutions
A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analysed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.

@@

% RAC
\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances organophosphate insecticides. 
Our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here \todo{zahlen neonics} and shows that this class may pose a high risk to freshwaters \todo{anderes Wort}.
Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality. 
This is also reflected in the different sample sizes (1,566 vs. 2,918,604 measurements \todo{use number for substances with RAC}) and the high number of non-detects in the monitoring data.

@@

% TU
xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshhold above biological effects have been shown \todo{add citation schaefer}. 
The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with one dominant compound.
% Mixtures
Nevertheless, most pesticides did not occur individually but in mixtures.
This highlights that
